---
title: "Unit 2"
output: html_document
date: "2023-09-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
 
 
The codes below requires the R package ISLR. Use install.packages("ISLR") to install the R package. Then load the R package.

```{r}
library(ISLR)
```



## Import and examine the data 

The dataset we will work with is called $Advertising$. It is available online and it is a csv file. You can view the dataset once imported. Since the first column $X$ is the index of the data points, it is removed. The last column $sales$ is the response column. 

```{r}
Advertising= read.csv("https://www.statlearning.com/s/Advertising.csv")
Advertising
Advertising$X = NULL
```


You can find the names of columns of the dataset. 
```{r}
names(Advertising)

```

A pairwise plot is useful to graphically look at the relationship between any two variables (including the response variable and the input varibles)

```{r}
pairs(Advertising)
```



## Model fitting
We first fit a linear regression model on $sales$ using the three independent variables. 

```{r}
fit = lm(sales~TV+radio+newspaper, data = Advertising)
fit = lm(sales~., data = Advertising)
```

A summary of the fitted model is obtained. 
```{r}
sfit = summary(fit)
sfit
str(sfit)
```

If we are interested in whether or not there is a relationship between advertising sales and budget, we are performing a hypothesis testing with the null hypothesis that there is no relationship between the   $sales$ and the indepedent variables. 
```{r}
sfit$fstatistic[1]
1- pf(sfit$fstatistic[1],sfit$fstatistic[2],sfit$fstatistic[3])
```
The $p$-value is extremely small. It implies that we will reject the null hypothesis and conclude that one or more independent variables affect the responses significantly. 


To answer how strong the relationship is, we check the R square. An R square value close to 1 indicates that the model explains a large portion of the variance in the response variable. 
```{r}
sfit$r.squared
```

We can check which media contributes to sales by looking at the summary table. 
```{r}
sfit
```

We can also check how large  the effect of each medium on sales is by computing the estimated regression coefficients and the confidence interval of the estimates. 
```{r}
sfit$coefficients
coef(sfit)
cbind(sfit$coefficients[,1]-qt(0.975,196)*sfit$coefficients[,2], sfit$coefficients[,1]+ qt(0.975,196)*sfit$coefficients[,2])
confint(fit)
```


We can make prediction at the observed and unobserved inputs. There are two different predictions, predicting the mean response and the individual response. 
```{r}
fitted.values = fitted(fit)
pred_clim = predict.lm(fit,data = Advertising,se.fit=TRUE, interval =  "confidence")
pred_plim = predict.lm(fit,data = Advertising,se.fit=TRUE, interval =  "prediction")
new = data.frame(TV = 60, radio = 6, newspaper = 26 )
pred_new_clim = predict.lm(fit, new,se.fit=TRUE, interval =  "confidence")
pred_new_plim = predict.lm(fit, new,se.fit=TRUE, interval =  "prediction")
```

Before making conclusion, we shall conduct model diagnostics. Verify the model assumptions. 
```{r}
par(mfrow=c(2,3))
plot(fitted.values, sfit$residuals,xlab = "fitted values", ylab="residuals")
plot(Advertising$TV, sfit$residuals,xlab = "TV", ylab="residuals")
plot(Advertising$radio, sfit$residuals,xlab = "radio", ylab="residuals")
plot(Advertising$newspaper, sfit$residuals,xlab = "newspaper", ylab="residuals")
plot(1:200, sfit$residuals,xlab = "Index",ylab = "residuals")
qqnorm(sfit$residuals)
qqline(sfit$residuals)
plot(fit)
```

We can also compare the full model with a smaller model or other models including the higher-order interaction terms. 
```{r}
fit2 = lm(sales~TV+radio, data = Advertising)
anova(fit2,fit)
```

The fitted model with the interaction is, 
```{r}
fit3 = lm(sales~TV*radio, data = Advertising)
anova(fit2,fit3)
summary(fit3)
```



 



 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
  
