---
title: "Unit 3"
output: html_document
date: "2023-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
### The probablity in logitistic regression

To understand the logistic regression, we make a  plot of  probability versus $x$. 

```{r}
beta0 = -13.5
beta1 = 12.5
x = seq(-5,10,by = 0.01)
plot(x, exp(beta0+beta1*x)/(1+exp(beta0+beta1*x)),xlab="x", ylab="probability")
```

One can make change the values of $\beta_0$ and $\beta_1$ and see how the probability changes accordingly. 

 


### Logistic regression for the Default data

The codes below requires the R package ISLR. Use install.packages("ISLR") to install the R package. Then load the R package.

```{r}
library(ISLR)
```


Read in the build-in data set *Default*. Use *str* to see the data structure.


```{r} 
data("Default")
str(Default)
```

Before fitting, we have a look at the data.

```{r}
plot(Default$balance[Default$default=="No"],Default$income[Default$default=="No"],type='p',col="blue",xlab="balance",ylab="income")
points(Default$balance[Default$default=="Yes"],Default$income[Default$default=="Yes"], col="red" )
boxplot(Default$balance~Default$default,xlab="default",ylab="balance")
boxplot(Default$income~Default$default,xlab="default",ylab="income")
```

We fit the logistic regression model either using the *glm* function or the *multinom* function.

```{r} 
model_logit = glm(default ~ student + balance + income, family = binomial(link='logit'), data=Default)
require(nnet)
model_logit2 <- multinom(default ~.,data=Default)
```

We can see the model outputs.
```{r}
summary(model_logit)
str(model_logit)
model_logit$coef
fit = model_logit$fit
#as.numeric(model_logit$fit>0.5)


summary(model_logit2)
str(model_logit2)
#model_logit2
#model_logit2$fit
#as.numeric(model_logit2$fit>0.5)
```

One can check the different aspects of the model fitting such as fitted values and residuals.

```{r}
#residuals(model_logit)
#fitted(model_logit)
coef(model_logit)
deviance(model_logit)
```


We can estimate a response as follows. For example, the estimated response at the first input is as,
```{r}
sum(c(1, 0, 729.5265, 44361.6251)*model_logit$coef)
exp(-6.549544)/(1+exp(-6.549544))
model_logit$fit[1]
```
 
We can also make prediction. For example, 

```{r}
sum(c(1, 0, 1500, 40000)*model_logit$coef)
exp(-2.142949)/(1+exp(-2.142949))

sum(c(1, 1, 1500, 40000)*model_logit$coef)
exp(-2.789725)/(1+exp(-2.789725))
```


We can use the *predict* function to make the prediction 
```{r}
Test_default = Default[1:50,2:4]
pred_logit= predict(model_logit,Test_default, type="response")
as.numeric(pred_logit>=0.5)
pred_logit2= predict(model_logit2,Test_default)
```

 

###  Linear Discriminant Analaysis 


To conduct linear discriminant analysis, we use the *lda* function in the R package *MASS*. 
 
```{r}
library(MASS)
model_lda = lda(default~ . ,data=Default)
model_lda
str(model_lda)
```

 

The following code is to help understand the estimate $\hat{\mu}$ in LDA. 
```{r}
mean(Default$balance)

mean(Default$balance[Default$default=='No'])
mean(Default$balance[Default$default=='Yes'])

mean(as.numeric(Default$student[Default$default=='No'])-1)
mean(as.numeric(Default$student[Default$default=='Yes'])-1)
```


We can use the *prediction* function to make prediction using LDA.

```{r}
pred_lda = predict(model_lda,Test_default)
pred_lda$class
pred_lda$posterior 
```
 
###  Quadratic Discriminant Analysis


We can perform quadratic discriminant analysis using the *qda* function. 

```{r}
model_qda = qda(default~ . ,data=Default)
model_qda
```

We can make prediction using the *predict* function.

```{r}
pred_qda = predict(model_qda,Test_default)
pred_qda$class
pred_qda$posterior 
```

 
### Comparison of the logistic regression, LDA and QDA.

To compare the logistic regression, LDA and QDA, we compute the confusion matrix. 

```{r}
# logitistic regression
pred_logit= predict(model_logit, Default, type="response")
pred_logit = as.numeric(pred_logit>=0.5)
cm_logit= table(pred_logit, Default$default)

# lda 
pred_lda = predict(model_lda,Default)
cm_lda = table(pred_lda$class, Default$default)


# qda 
pred_qda = predict(model_qda,Default)
cm_qda = table(pred_qda$class, Default$default)
```


Compare the different model using *precision*.
```{r}
precison_logit = cm_logit[2,2]/(cm_logit[2,1] + cm_logit[2,2])
precison_lda = cm_lda[2,2]/(cm_lda[2,1] + cm_lda[2,2])
precison_qda = cm_qda[2,2]/(cm_qda[2,1] + cm_qda[2,2])
c(precison_logit,precison_lda,precison_qda)
```

Compare the different model using *sensitivity*. 

```{r}
sensitivity_logit = cm_logit[2,2]/(cm_logit[1,2] + cm_logit[2,2])
sensitivity_lda = cm_lda[2,2]/(cm_lda[1,2] + cm_lda[2,2])
sensitivity_qda = cm_qda[2,2]/(cm_qda[1,2] + cm_qda[2,2])
c(sensitivity_logit,sensitivity_lda,sensitivity_qda)
```

Compare the different model using *Specifity*. 

```{r}
specifity_logit = cm_logit[1,1]/(cm_logit[1,1] + cm_logit[2,1])
specifity_lda = cm_lda[1,1]/(cm_lda[1,1] + cm_lda[2,1])
specifity_qda = cm_qda[1,1]/(cm_qda[1,1] + cm_qda[2,1])
c(specifity_logit,specifity_lda,specifity_qda)
```

 

 
 


 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
  
